# Reproducible Research Assignment 1

```{r results="hide"}
library(ggplot2)
library(Hmisc)
```

## Loading Dataset 

```{r}
Actdata<-read.csv("activity.csv")
head(Actdata)
```

## Q1-What is the mean total number of steps taken per day

### 1.1 Calculate total number of steps taken per day

```{r}
stepsperday<-tapply(Actdata$steps,Actdata$date,sum,na.rm=TRUE)
```

### 1.2 Histogram of total number of steps taken each day

```{r}
hist(stepsperday,xlab = "Number of Steps", ylab = "Number of Days", main="Histogram of Steps/Day", breaks = 10 )
```

### 1.3 Mean and Median of the Total Number of Steps Taken per day

```{r}
meanstepsperday<-mean(stepsperday)
medianstepsperday<-median(stepsperday)
print(meanstepsperday)
print(medianstepsperday)
```

## Q2-What is the Average daily activity pattern?

### 2.1 Time Series Plot of 5 min interval vs average number of steps taken accross all days

```{r}
stepsperinterval<-aggregate(list(steps=Actdata$steps),list(interval=Actdata$interval),mean,na.rm=TRUE)
plot(stepsperinterval,type="l", xlab="Interval", ylab="Number of Steps", main="Time Series Plot")
```

### 2.2 5 minute interval that contains maximum steps averaged accross all days

```{r}
maxstepsinterval<-stepsperinterval[which.max(stepsperinterval$steps),]
print(maxstepsinterval)
```

## Q3- Input Missing values in the data

### 3.1 Calculate and report the total number of missing values in the dataset 

```{r}
missingdata<-sum(is.na(Actdata$steps))
print(missingdata)
```

### 3.2 New dataset with imputed missing values

```{r}
newActdata<-Actdata
newActdata$steps<-impute(Actdata$steps,mean)
head(newActdata)
```

### 3.3 Histogram of total number of steps taken per day and updated mean & median
```{r}
newstepsperday<-tapply(newActdata$steps,newActdata$date,sum)
hist(newstepsperday,xlab = "Number of Steps", ylab = "Number of Days", main="Histogram of Steps/Day (After Imputing Missing Values)")
newmeanstepsperday<-mean(newstepsperday)
newmedianstepsperday<-median(newstepsperday)
print(newmeanstepsperday)
print(newmedianstepsperday)
```

#### NoTE: We do see and increase in the mean and the median once we fill the missing values

## Q4- Are there differences in activity patterns between weekdays and weekends?

### 4.1  a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
newActdata$typeofday<-ifelse(as.POSIXlt(newActdata$date)$wday %in% c(0,6),"weekend","weekday")
head(newActdata)
```

### 4.2 Panel time series plot comparing weekdays and weekends
```{r}
newstepsperinterval<-aggregate(list(steps=newActdata$steps),list(interval=newActdata$interval,typeofday=newActdata$typeofday),mean,na.rm=TRUE)
ggplot(newstepsperinterval, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(typeofday ~ .) +
  xlab("5-minute interval") + 
  ylab("avarage number of steps")
```



